---
title: "Practical Machine Learning Course Project"
author: "Siti Nurillah Binti Mohd Fadzil"
date: "January 29, 2016"
output: html_document
---

##Data 

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


##Reading and Cleaning Data

Start with reading both training and testing instances. 


```{r}
library(data.table)
library(caret)
library(randomForest)
library(foreach)
library(rpart)
library(rpart.plot)
library(corrplot)
library(ggplot2)
```

```{r}
#read training data
training_data <- read.csv("./pml-training.csv", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))
#read test data
testing_data <- read.csv("./pml-testing.csv", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))
```

Columns with NAs will be removed. Besides that highly correlated variables and variables with 0 (or approx to 0) variance will be removed. 

#Cleaning data

```{r}
clean_training <- training_data[, -which(names(training_data) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window"))]

#remove columns with NAs
clean_training = clean_training[, colSums(is.na(clean_training)) == 0]

#remove variables with 0 or near to 0 variance
zero_variance =nearZeroVar(clean_training[sapply(clean_training, is.numeric)], saveMetrics=TRUE)
clean_training = clean_training[, zero_variance[, 'nzv'] == 0]

correlation_matrix <- cor(na.omit(clean_training[sapply(clean_training, is.numeric)]))

dim(correlation_matrix)
correlationmatrixdegreesoffreedom <- expand.grid(row = 1:52, col = 1:52)
 #this returns the correlation matrix in matrix format
correlationmatrixdegreesoffreedom$correlation <- as.vector(correlation_matrix)
removehighcorrelation <- findCorrelation(correlation_matrix, cutoff = .7, verbose = TRUE)
#this removes highly correlated variables (in psychometric theory .7+ correlation is a high correlation)
clean_training <- clean_training[, -removehighcorrelation] 

```


```{r}

for(i in c(8:ncol(clean_training)-1)) {clean_training[,i] = as.numeric(as.character(clean_training[,i]))}

for(i in c(8:ncol(testing_data)-1)) {testing_data[,i] = as.numeric(as.character(testing_data[,i]))}
```

The cleaned dataset will only consist of complete columns. For a lighter dataset, user name, timestamps and windows will be removed.

```{r}
#drop blank column
featureset <- colnames(clean_training[colSums(is.na(clean_training)) == 0])[-(1:7)]
modeldata <- clean_training[featureset]

#cleansed data to build model
featureset
```



##Cross-Validation
#Split the sample in two samples. This is to divide training and testing for cross-validation.

```{r}
idx <- createDataPartition(modeldata$classe, p=0.6, list=FALSE )
training <- modeldata[idx,]
testing <- modeldata[-idx,]

control <- trainControl(method="cv", 5)
model <- train(classe ~ ., data=training, method="rf", trControl=control, ntree=250)
model

predict <- predict(model, testing)
confusionMatrix(testing$classe, predict)

accuracy <- postResample(predict, testing$classe)
accuracy

result <- predict(model, training[, -length(names(training))])
result

treeModel <- rpart(classe ~ ., data=clean_training, method="class")
prp(treeModel) 
```

##Conclusions and Test Data Submit

As can be seen from the confusion matrix the proposed model is very accurate. 

Prepare the submission. (using COURSERA provided code)

```{r, eval=FALSE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```